---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, setting_options}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

Factor analysis visualization made easy with FAtools
=================
[![Build Status](https://travis-ci.org/mattkcole/FAtools.svg?branch=master)](https://travis-ci.org/mattkcole/FAtools)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/mattkcole/FAtools?branch=master&svg=true)](https://ci.appveyor.com/project/mattkcole/FAtools)
[![Coverage Status](https://img.shields.io/codecov/c/github/mattkcole/FAtools/master.svg)](https://codecov.io/github/mattkcole/FAtools?branch=master)


### NOTE: THIS PACKAGE IS IN DEVELOPMENT

From choosing the numbers of factors to extract to inspecting loadings, factor analysis can be very visual in nature.
The FAtools R package aims to make this process easier by providing functions to do visualizations with ease.

### To Download:

```{r message=FALSE, loadingpackage, warning=FALSE}
library('devtools')
#devtools::install_github('mattkcole/FAtools')
library('FAtools')
```

### Examples:

We can first look at our data (here we are using the possibly cliche but familiar data, mtcars).
```{r loadingdata}
library(datasets)
summary(mtcars)
```

Let's first make our correlation matrix - we wont worry about scaling or investigating our data much for this demonstration (usually a bad idea).
```{r gettingcor}
corr.matrix <- cor(mtcars)
```

Let's load the packages we need for our analysis:
```{r lodingpackages2, cache=F, message=FALSE, warning=FALSE}
library('psych')    # for statistical methods
library('FAtools')  # for some plotting and EDA
library('dplyr')    # for data wrangling
library('knitr')    # for rmd help
```

Lets make and plot our scree plot to assess the number of factors present.
```{r scree_plot, cache=TRUE}
s.plot <- FAtools::scree_plot(corr.matrix, nrow(mtcars), ncol(mtcars))
plot(s.plot)
```

We can conduct our factor analysis with two factors using the psych package.
```{r loadings}
results <- psych::fa(corr.matrix, 2, rotate = "varimax")
results$loadings
```


The loadings look pretty good, but we can make them more interpretable by excluding low loadings (param: `cutoff`), rounding (param: `roundto`), incorporate a data dictionary, and include labels -- And we can use the knitr::kable() function for great looking tables in Rmarkdown documents.  

```{r}
FAtools::loadings_table(results$loadings, 2, cutoff = 0.3, roundto = 2) %>%
        kable()
```

Say we had more informative names than `colnames(mtcars)`.

```{r}
cool_names <- c("Miles Per Gallon", "Cylinders", "Displacement",
                "Gross horsepower", "Rear Axle ratio", "Weight (1K lbs)",
                "1/4 mile time", "V/S", "Manual", "Forward gears",
                "Carburetors")
```

And say we wern't really all that interested in loadings with an absolute value less than 0.3. 

```{r }
FAtools::loadings_table(loading_frame = results$loadings,
                        cutoff = 0.3, roundto = 2,
                        Name = colnames(mtcars), 
                        description = cool_names) %>%
        kable()
```


We could also display this graphically, which works well when we have more retained factors or many more variables. (let's say we have 5 factors extracted).

```{r factor_plot, fig.width=8, fig.height=4.5, cache=TRUE}
loadings5 <- cor(mtcars) %>%
        psych::fa(2, rotate = "varimax")
        
FAtools::loadings_plot(loadings5$loadings,
                       colorbreaks = c(0, 0.2, 0.4, 0.6, 0.8, 1),
                       labRow = c("F1", "F2"),
                       columnlabels = cool_names)
```

Looks great! 



Submit and issue with any concerns! 



Credits:
Much of the scree plot functionality comes from code provided by:
[www.statmethods.net](http://www.statmethods.net/advstats/)

