---
title: "readme.rmd"
author: "Matt Cole"
date: "2/18/2017"
output:
  md_document:
    variant: markdown_github
---
Factor analysis visualization made easy with FAtools
=================
[![Build Status](https://travis-ci.org/mattkcole/FAtools.svg?branch=master)](https://travis-ci.org/mattkcole/FAtools)


From choosing the numbers of factors to extract to inspecting loadings, factor analysis can be very visual in nature.
The FAtools R package aims to make this process easier by providing functions to do visualizations with ease.

### To Download:

```{r message=FALSE}
library('devtools')
devtools::install_github('mattkcole/FAtools')
```

### Examples:

We can first look at our data (here we are using the possibly cliche but familiar data, mtcars).
```{r}
library(datasets)
summary(mtcars)
```

Let's first make our correlation matrix - we wont worry about scaling or investigating our data much for this demonstration (usually a bad idea).
```{r}
corr.matrix <- cor(mtcars)
```

Let's load the packages we need for our analysis:
```{r}
library('psych')
library('FAtools')
library('MASS')
```

Lets make and plot our scree plot to assess the number of factors present.
```{r}
# s.plot <- FAtools::scree_plot(corr.matrix, nrow(mtcars), ncol(mtcars))
# plot(s.plot)
```

We can conduct our factor analysis with two factors using the psych package.
```{r}
results <- psych::fa(corr.matrix, 2)
results$loadings
```


The loadings look pretty good, but we can make them more interpretable by excluding low loadings (param: `cutoff`), rounding (param: `roundto`), incorporate a data dictionary, and include labels -- And we can use the knitr::kable() function for great looking tables in Rmarkdown documents.  

```{r}
library(knitr)
```

```{r}
FAtools::loadings_table(results$loadings, 2, cutoff = 0.3, roundto = 2)
```

Say we had more informative names than `colnames(mtcars)`.

```{r}
cool_names <- c("Miles Per Gallon", "Cylinders", "Displacement",
                "Gross horsepower", "Rear Axle ratio", "Weight (1K lbs)",
                "1/4 mile time", "V/S", "Manual", "Number forward gears",
                "Number of carburetors")
```

And say we wern't really all that interested in loadings with an absolute value less than 0.3. 

```{r}
FAtools::loadings_table(loading_frame = results$loadings, loadings_no = 2,
                        cutoff = 0.3, roundto = 2,
                        Name = colnames(mtcars), 
                        Description = cool_names)
```


We could also display this graphically, which works well when we have more retained factors or many more variables.

```{r}
FAtools::loadings_plot(loadings = results$loadings,
                       cool_names,
                       colorbreaks = c(-.2,0.4,0.6,0.8,1),
                       colors = RColorBrewer::brewer.pal(4,"Greens"))
```


Submit and issue with any concerns! 



Credits:
Much of the scree plot functionality comes from code provided by:
[www.statmethods.net](http://www.statmethods.net/advstats/factor.html)

